---
title: "R Notebook"
---

Get phyloseq object ready for metadata
* get sample names to match metadata
* will need to duplicate metadata rows for duplicate samples
* collect extraction + library prep sample info (qubit (pre-PCR, pre-pooling), PCR fail?, possible contamination, extraction notes)
* keep sample name and ring number where possible to catch mislabelled tubes or DB errors by comparing to birdData info sheet
  * need to match sample ID and bird ID from SummerSamples.csv sheet to sampleData_Lab sheet

```{r}
rm(list=ls())
```

```{r}
library(tidyverse)
```


```{r}
sampleData_LibraryPrep <- read.csv2("../Data/Dataframes_raw_csv/sampleData_LibraryPrep.csv", header = TRUE, sep = ",")
SummerSamples <- read.csv2("../Data/Dataframes_raw_csv/SummerSamples.csv", sep = ",")
```

Add new col to sampleData of ring number
* merge ring number from SummerSamples
*need to verify these all correctly merged | any rows duplicated*
  * looks like 4 extra rows
```{r}
sampleData_Lab <- merge(sampleData_LibraryPrep, SummerSamples[,c("Bird_ID", "Sample_ID")], by = "Sample_ID", all.x = TRUE)
```

Multiple rings matching an ID
* CB23.Fe [note: 2 different ring numbers] PF91038(matches individual_DF), PF91049(matches CB56.Fe in individual_DF)
* CB62.D.D15 [note: 2 different ring numbers] TX52575(matches individual_DF), PF91014(matches CB62.F in ind_DF)
* KB22.A.D15 [note: 2 different ring numbers] TX52580(matches indv.DF) TX52581(matches KB22.B in indiv_DF)

* KB908.B.D8 [contaminated duplicate] has 4 copies because 2 copies in summerSamples and 2 copies in sampleData_libraryprep (i think)
```{r}
which(duplicated(sampleData_Lab$Azenta_noPeriod))
```

Change summer samples sample ID according to bird ID
Change sampleData_library sample ID according to azenta_ID

```{r}
SummerSamples_fix <- SummerSamples

SummerSamples_fix[SummerSamples_fix$Bird_ID=="PF91049", "Sample_ID"] <- "CB56.Fe" # missing CB56.Fe sample data so confident this is correct sample, '49 is CB56.Fe according to birdData.csv
SummerSamples_fix[SummerSamples_fix$Bird_ID=="PF91014", "Sample_ID"] <- "CB62.F.D15" # no D8 data collected for CB62.F so confident its D15 sample
SummerSamples_fix[SummerSamples_fix$Bird_ID=="TX52581" & SummerSamples_fix$Sample_ID=="KB22.A.D15", "Sample_ID"] <- "KB22.B.D15" # have D15 data for A.D15 and B.D8 already so confident this is B.D15
```

```{r}
sampleData_LibraryPrep_fix <- sampleData_LibraryPrep

sampleData_LibraryPrep_fix[sampleData_LibraryPrep_fix$Azenta_ID=="CB23.Fe.2", "Sample_ID"] <- "CB56.Fe" #'49 extracted second and hence later in library prep so changing '.2 rather than '.1
sampleData_LibraryPrep_fix[sampleData_LibraryPrep_fix$Azenta_noPeriod=="CB62xDxD15xP3", "Sample_ID"] <- "CB62.F.D15"

sampleData_LibraryPrep_fix[sampleData_LibraryPrep_fix$Azenta_noPeriod=="KB22xAxD15xP2", "Sample_ID"] <- "KB22.B.D15" # 'xP2 extraction dates match up with '81 and was extracted before 'xP3/'80
```

Label contaminated version of KB908.B.D8 (sample interrupted by fire alarm one was contam in field)
* Azenta_ID = KB908.B.D8.2 is contaminated
* second one to be extracted was fire alarm one hence extraction number 26 rather than 13
* why are these being duplicated? Because kb908.b is the only duplicate sample name in summerSamples
* temporarily make it unique. Then merge. then return it to original sample name
```{r}
SummerSamples_fix[SummerSamples_fix$Extraction_number==26, "Sample_ID"] <- "KB908.B.D8.contam"
 
sampleData_LibraryPrep_fix[sampleData_LibraryPrep_fix$Azenta_ID=="KB908.B.D8.2", "Sample_ID"] <- "KB908.B.D8.contam"
```

Re-merge dataframes without incorrect Sample IDs
```{r}
sampleData_Lab_fix <- merge(sampleData_LibraryPrep_fix, SummerSamples_fix[,c("Bird_ID", "Sample_ID")], by = "Sample_ID", all = T)

# verify no Azenta duplicates
which(duplicated(sampleData_Lab_fix$Azenta_noPeriod))

# return KB908.B to original sample name post merge so it can match to bird data
sampleData_Lab_fix[sampleData_Lab_fix$Azenta_ID=="KB908.B.D8.2", "Sample_ID"] <- "KB908.B.D8"
```

KB10 sample was mislabelled as '.F' but probably was KB10.E.D15
```{r}
sampleData_Lab_fix[sampleData_Lab_fix$Sample_ID=="KB10.F", "Sample_ID"] <- "KB10.E.D15"
# add note
sampleData_Lab_fix[sampleData_Lab_fix$Sample_ID=="KB10.E.D15", "Post_lab_notes"] <- "KB10 sample was mislabelled as '.F' but probably was KB10.E.D15"
```

Sample Cb20.Fe is actually KB20.Fe, there was no CB20 nest in the study and ring number matches KB20.
```{r}
sampleData_Lab_fix[sampleData_Lab_fix$Sample_ID=="CB20.Fe", "Sample_ID"] <- "KB20.Fe"
# add note
sampleData_Lab_fix[sampleData_Lab_fix$Sample_ID=="KB10.E.D15", "Post_lab_notes"] <- "KB20.Fe sample was mislabelled as 'CB20.fe' on tube"
```

Sample "CB26.X.D15?", i am unsure of chick IDletter, either C or D
* drop '?' as it is stopping merge of bird data
```{r}
sampleData_Lab_fix[sampleData_Lab_fix$Sample_ID=="CB26.X.D15?", "Sample_ID"] <- "CB26.X.D15"
```

Correct wrong ring numbers
```{r}
# KB15.M: TW15318, TW15218
sampleData_Lab_fix[sampleData_Lab_fix$Azenta_noPeriod =="KB15xMxP1", "Bird_ID"] <- "TW15318"

# KB22.Fe: TX52595, TX52594
sampleData_Lab_fix[sampleData_Lab_fix$Azenta_noPeriod=="KB22xFexP1", "Bird_ID"] <- "TX52595"

# Note ring number disagrees between KB29.M.1 and M.2
sampleData_Lab_fix[sampleData_Lab_fix$Sample_ID=="KB29.M.2", "Bird_ID"] <- "TX52594"

```


```{r}
write_csv(sampleData_Lab_fix, file = "../Data/Dataframes_R_out/LabSampleData.csv")
```

