---
title: "R Notebook"
---

Script to plot and test variance of alpha diversity, and how it is affected by treatment

```{r, echo=FALSE}
rm(list=ls())
```

# libraries
```{r, , echo=FALSE, message=FALSE}
library(tidyverse)
library(phyloseq)
library(microbiome)
library(kableExtra)
```

# data read-in
```{r, echo=FALSE}
ps <- readRDS("../Data/Phylo_objects/phylo-experimental-decontamPrevalence.rds")
metadata <- microbiome::meta(ps)
```

# Chao1 variance
## Plot variance

Overall age groups plot by treatment
```{r}
chao.violin <- ggplot(metadata, aes(Treatment, log(Chao1))) +
  geom_violin() +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  theme_classic()

chao.violin
```


Test for homogeneity of variance among treatment groups
```{r}
bartlett.test(log(metadata$Chao1), metadata$Treatment)
```

# Shannon variance
## Plot variance

Overall age groups plot by treatment
```{r}
shannon.violin <- ggplot(metadata, aes(Treatment, Shannon)) +
  geom_violin() +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  theme_classic()

shannon.violin
```

Test for homogeneity of variance among treatment groups
```{r}
bartlett.test(metadata$Shannon, metadata$Treatment)
```

# PD variance
## Plot variance

Overall age groups plot by treatment
```{r}
PD.violin <- ggplot(metadata, aes(Treatment, log(PD))) +
  geom_violin() +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  theme_classic()

PD.violin
```

Test for homogeneity of variance among treatment groups
```{r}
bartlett.test(log(metadata$PD), metadata$Treatment)
```