---
title: "R Notebook"
---

Look at relative abundance of firmicutes to see if diversity effect of treatment is being driven by increase in firmicutes

```{r}
rm(list=ls())
```

```{r}
library(phyloseq)
library(microbiome)
library(tidyverse)
library(lme4)
library(wec)
library(DHARMa)
library(kableExtra)
```

```{r}
ps <- readRDS("../Data/Phylo_objects/phylo-experimental-decontamPrevalence.rds")

metadata <- microbiome::meta(ps)
```

Recalculate number of reads following decontam
```{r}
metadata$NumberOfReads <- sample_sums(ps)
ps@sam_data$NumberOfReads <- sample_sums(ps)
```

Create and save a melted df of data glommed to phyla level
```{r}
#sample_data(phylo.spring) <- metadata.scaled
ps.glom <-  tax_glom(ps, "Phylum", NArm = TRUE)
mot <- psmelt(ps.glom)

```

```{r}
firmicutes.df <- mot[mot$Phylum=="Firmicutes",]
```

```{r, echo=FALSE}
# set age as factor
firmicutes.df$Age.category <- factor(firmicutes.df$Age.category, levels = c("D8", "D15", "Adult"))
```

## WEC
Create weighted contrast encoding, specifying "Adult" as omitted category in first variable and D8 omitted in second variable
```{r}
firmicutes.df$Age.category.wec1 <- firmicutes.df$Age.category
contrasts(firmicutes.df$Age.category.wec1) <- contr.wec(firmicutes.df$Age.category, "Adult")
contrasts(firmicutes.df$Age.category.wec1)

firmicutes.df$Age.category.wec2 <- firmicutes.df$Age.category
contrasts(firmicutes.df$Age.category.wec2) <- contr.wec(firmicutes.df$Age.category, "D8")
contrasts(firmicutes.df$Age.category.wec2)
```

# Model

## dummy coded model
```{r}
binomial.response <- c(firmicutes.df$Abundance/firmicutes.df$NumberOfReads)

bin.firmicutes <- glmer(binomial.response ~ Treatment + Age.category + (1|Site/Nest/Bird.ID), family= binomial, weights = NumberOfReads, data = firmicutes.df)
summary(bin.firmicutes)
```

## wec model
```{r}
binomial.response <- c(firmicutes.df$Abundance/firmicutes.df$NumberOfReads)

bin.firmicutes.wec1 <- glmer(binomial.response ~ Treatment + Age.category.wec1 + (1|Site/Nest/Bird.ID), family= binomial, weights = NumberOfReads, data = firmicutes.df)
summary(bin.firmicutes.wec1)
```

```{r}
binomial.response <- c(firmicutes.df$Abundance/firmicutes.df$NumberOfReads)

bin.firmicutes.wec2 <- glmer(binomial.response ~ Treatment + Age.category.wec2 + (1|Site/Nest/Bird.ID), family= binomial, weights = NumberOfReads, data = firmicutes.df)
summary(bin.firmicutes.wec2)
```


# Check models

```{r}
simulationOutput <- simulateResiduals(fittedModel = bin.firmicutes, n = 250)
plot(simulationOutput)
```

```{r}
simulationOutput <- simulateResiduals(fittedModel = bin.firmicutes.wec1, n = 250)
plot(simulationOutput)
```

```{r}
simulationOutput <- simulateResiduals(fittedModel = bin.firmicutes.wec2, n = 250)
plot(simulationOutput)
```

# results

Use wec models
```{r}
firmic.wec1.coef <- summary(bin.firmicutes.wec1)$coefficients # returns summary table for wec1 model omitting adult
firmic.wec2.coef <- summary(bin.firmicutes.wec2)$coefficients["Age.category.wec2Adult",] #gets adult estimates
# 
firmic.coef <- rbind(firmic.wec1.coef, firmic.wec2.coef)
# 
rownames(firmic.coef) <- c("(Intercept)", "Treatment (L. kimchicus)", "Age (D8)","Age (D15)", "Age (Adult)")

#make df of lmer output
firmic.coef.df <- as_tibble(firmic.coef, rownames="Independent variables") %>%
  dplyr::rename("p"="Pr(>|z|)") %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(p=ifelse(p==0,"<0.001",p)) %>% 
  mutate(p=ifelse(p<=0.05,str_c(p," *"),p)) %>%
  mutate(p=ifelse(p>=0.05 & p<=0.06,str_c(p,"  ."),p))


# make df into kable
## this is in html, dosnt render in word doc # but can copy-paste
results.table <- kable(firmic.coef.df, format = "html", table.attr = "style = \"color: black;\"") %>%
  kableExtra::kable_styling(full_width = F) #%>%
  #save_kable("alpha-kable__________.png") # this line saves as .png in Reports/

results.table
```

