---
title: "R Notebook"
---

How does L. kimchicus treatment affect the microbiome: plots

```{r}
rm(list=ls())
```

```{r}
#devtools::install_github("thomazbastiaanssen/Tjazi")

library(tidyverse)
library(phyloseq)
library(microbiome)
library(Tjazi)
library(ggbeeswarm)
library(metafolio)
library(compositions)
library(vegan)
library(cowplot)
```

# read in data

```{r}
ps <- readRDS("../Data/Phylo_objects/phylo-experimental-decontamPrevalence.rds")
metadata <- microbiome::meta(ps)

metadata$Age.category    = factor(metadata$Age.category, levels = c("D8", "D15", "Adult"))
```

# Raw plots
# alpha diversity

Simply use metadata alpha div for plotting
```{r}
alpha_diversity <- metadata[,c("Sample.ID", "Treatment", "SampleType", "Shannon", "Chao1", "PD")]
```

Change treatment level L kimchicus to 'Treatment'
```{r}
alpha_diversity$Treatment <- ifelse(alpha_diversity$Treatment == "L. kimchicus", "Treatment", "Control")
```

Log chao and Faiths PD
```{r}
alpha_diversity$log.chao <- log(alpha_diversity$Chao1)
alpha_diversity$log.PD <- log(alpha_diversity$PD)

alpha_diversity <- alpha_diversity[,!colnames(alpha_diversity) %in% c("Chao1")] %>%
   dplyr::rename("log(Chao1)" = "log.chao") # drop raw Chao1 in favour of log chao1

alpha_diversity <- alpha_diversity[,!colnames(alpha_diversity) %in% c("PD")] %>%
   dplyr::rename("log(PD)" = "log.PD")
```


What is the weird extreme shannon entropy outlier, it has also moved from control to probiotic here compared to 06_ script?
* highest diversity bird in metadata is KB25xCxD8xP3 a 'probiotic' bird
* highest shannon entropy bird in alpha diversity is DW15xCxD15xP3
* there is no outlier in the metadata (i.e. not Tjazi calculated alpha div) it seems unless ive changed something since
```{r}

#Plot alpha diversity all at once using pipes
all.alpha.tjazi <- alpha_diversity %>%
#Wrangle the data to long format for easy plotting
  pivot_longer(!c(Treatment, SampleType, Sample.ID)) %>%
  #Pipe it all directly into ggplot2
  ggplot(aes(x     = Treatment,
             y     = value,
             fill  = Treatment,
             #shape = Sex,
             group = Treatment)) +
  geom_boxplot(alpha = 1/2, coef = 100) +
  geom_jitter(width = 0.2,size = 1, cex = 3, alpha = 0.35) +
  facet_wrap(~name, scales = "free") + theme_bw()  +
  scale_fill_manual(values = c("Control"        = "#49b7fc",
                                "Treatment" = "#ff7b00")) +
  theme(axis.title.x = element_blank(), legend.title = element_blank()) +
  ylab("Diversity") +
  xlab("") +
  theme_classic()

all.alpha.tjazi

ggsave("Figure1.pdf", plot = all.alpha.tjazi, device = "pdf", path = "../Figures/", width = 15, height = 10, units = "cm", dpi = 300)
```

# Beta diversity (PCA)

```{r}
ps <- readRDS("../Data/Phylo_objects/phylo-experimental-decontamPrevalence.rds")
```

## No repeats
```{r}
set.seed(1189)
metadata.NoRepeats <- metadata %>% 
    group_by(Bird.ID) %>%
    sample_n(1)

dropped.samples <- setdiff(metadata$Sample.ID, metadata.NoRepeats$Sample.ID)

# metadata.even <- metadata %>% 
#     group_by(bird.ID) %>%
#     sample_n(min(table(metadata$bird.ID)))

table(metadata$Age.category)
table(metadata.NoRepeats$Age.category)

phylo.NoRepeats <- subset_samples(ps, Sample.ID %in%  metadata.NoRepeats$Sample.ID)

meta.noRepeat <- meta(phylo.NoRepeats)

# filter out rare taxa
ps.5pcFilt <- filter_taxa(phylo.NoRepeats, function(x) sum(x > 1) > (0.05*length(x)), TRUE)

# tjazi approach below
clr.asvTable <- t(Tjazi::clr_c(t(ps.5pcFilt@otu_table))) 

# getting 1 samples with all NAs, so i will simply drop those
clr.asvTable <- na.omit(clr.asvTable)
# calculate aitchison dist
aitchison.distances <- dist(clr.asvTable, method = "euclidean")
```

```{r}
# code from bugs as features
#Apply the base R principal component analysis function on our CLR-transformed data.
#data.a.pca  <- prcomp(t(genus.exp))
data.a.pca  <- prcomp(clr.asvTable)
#Extract the amount of variance the first four components explain for plotting.
pc1 <- round(data.a.pca$sdev[1]^2/sum(data.a.pca$sdev^2),4) * 100
pc2 <- round(data.a.pca$sdev[2]^2/sum(data.a.pca$sdev^2),4) * 100
#Extract the scores for every sample for the first four components for plotting.
pca  = data.frame(PC1 = data.a.pca$x[,1],
                  PC2 = data.a.pca$x[,2])

#subset metadata to just relevant samples, as dropped samples with NAs
meta.noRepeat.pca <- meta.noRepeat[rownames(meta.noRepeat) %in% rownames(clr.asvTable),]

#Add relevant information from the metadata
pca$Sample.ID        = meta.noRepeat.pca$Sample.ID
pca$Treatment        = meta.noRepeat.pca$Treatment
pca$Treatment <- ifelse(pca$Treatment == "L. kimchicus", "Treatment", "Control")
pca$Site             = meta.noRepeat.pca$Site
pca$Age.category = factor(meta.noRepeat.pca$Age.category, levels = c("D8", "D15", "Adult"))
```


```{r}
noRepeats_beta.correction  <- ggplot(pca, aes(x  = PC1,
                             y  = PC2,
                             fill= Treatment,
                             colour  = Treatment,
                             group   = Treatment)) +
  #Create the points and ellipses
  stat_ellipse(geom = "polygon", alpha = 1/4) +
  geom_point(size=3, colour = "black", pch=21) +
  #Adjust appearance
  scale_fill_manual(values   = c("Control" = "#49b7fc", "Treatment" = "#ff7b00")) +
  scale_colour_manual(values = c("Control" = "#49b7fc", "Treatment" = "#ff7b00")) +
  guides(fill = guide_legend(override.aes = list(shape = c(21)))) +
  #Adjust labels
  xlab(paste("PC1: ", pc1,  "%", sep="")) +
  ylab(paste("PC2: ", pc2,  "%", sep="")) +
  facet_grid(~Age.category) +
  theme_classic() +
  theme(legend.title = element_blank())

noRepeats_beta.correction

ggsave("Figure3.pdf", plot = noRepeats_beta.correction, device = "pdf", path = "../Figures/", width = 15, height = 10, units = "cm", dpi = 300)
```


Facet by site
* separate PCA for each site might be clearer
```{r}

noRepeats_beta.site <- 
noRepeats_beta.correction +
    #Adjust appearance
  scale_fill_manual(values   = c("Control" = "#49b7fc", "Treatment" = "#ff7b00")) +
  scale_colour_manual(values = c("Control" = "#49b7fc", "Treatment" = "#ff7b00")) +
  facet_wrap(~Site, nrow = 2)

noRepeats_beta.site

ggsave("Figure_S1.pdf", plot = noRepeats_beta.site, device = "pdf", path = "../Figures/", width = 15, height = 10, units = "cm", dpi = 300)
```
