---
title: "R Notebook"
---

Analysis of treatment effect on body weight

```{r}
rm(list=ls())
```

# libraries
```{r}
#install.packages("lme4")
library(lme4)
library(lmerTest)
library(tidyverse)

library(phyloseq)
library(microbiome)
library(kableExtra)
library(DHARMa)
```

# read in data
```{r}
ps <- readRDS("../Data/Phylo_objects/phylo-experimental-decontamPrevalence.rds")
metadata <- microbiome::meta(ps)
```

```{r}
#str(metadata)

# set age as factor
metadata$Age.category <- factor(metadata$Age.category, levels = c("D8", "D15", "Adult"))
```

## Backwards diff coding
```{r}
my.backward.diff = matrix(c(-2/3, 1/3, 1/3, -1/3, -1/3, 2/3), ncol = 2)
my.backward.diff
```


Add log.chao and PD variable
```{r}
metadata$log.chao <- log(metadata$Chao1)
metadata$log.PD <- log(metadata$PD)

```

## Standardize variables

```{r}
# centre and scale
numeric.predictors <- c("Wing_mm","Tarsus_mm","Weight_g", "Chao1", "Shannon", "PD", "BroodSize.D8", "BroodSize.D15", "firstEggLayDate", "lastEggLayDate", "log.chao", "log.PD") #Weight_g.scaled
metadata.scaled <- metadata
metadata.scaled[,numeric.predictors] <- scale(metadata.scaled[,numeric.predictors], scale = TRUE, center = TRUE)
```


```{r}
metadata.scaled$Age.category.BDC <- factor(metadata.scaled$Age.category, levels = c("D8", "D15", "Adult"))
levels(metadata.scaled$Age.category.BDC) # verify: 1week,2week,adult
#assigning the backward difference coding to ageBinned
contrasts(metadata.scaled$Age.category.BDC) = my.backward.diff
```

Create nestling only dataframe and scale separately
```{r}
metadata.nestling <- subset(x = metadata, subset = metadata$Age.category %in% c("D8", "D15"))
metadata.nestling[,numeric.predictors] <- scale(metadata.nestling[,numeric.predictors], scale = TRUE, center = TRUE)
```

# create paired dataframe

I will create the paired df from the original metadata and rescale separately
```{r}
# create D8 and D15 dfs
d8.df <- metadata[metadata$Age.category=="D8",c("Bird.ID","Sample.ID", "Wing_mm","Tarsus_mm","Weight_g", "Chao1", "Shannon", "log.chao", "PD", "log.PD")] #, "scaled.mass.d8", "Site","Nest","Treatment",
d15.df <- metadata[metadata$Age.category=="D15",c("Bird.ID","Sample.ID", "Wing_mm","Tarsus_mm","Weight_g", "Chao1", "Shannon", "PD", "log.PD")] # , "scaled.mass.d15"

# append .D8 and .D15 to appropriate columns
# rename columns with appendage
d8.df <- dplyr::rename(d8.df, Sample.ID.d8=Sample.ID, Wing_mm.d8=Wing_mm, Tarsus_mm.d8=Tarsus_mm, 
                Weight_g.d8=Weight_g, Chao1.d8=Chao1, Shannon.d8=Shannon, log.chao.d8=log.chao, PD.d8=PD, log.PD.d8=log.PD)
d15.df <- dplyr::rename(d15.df, Sample.ID.d15=Sample.ID, Wing_mm.d15=Wing_mm, Tarsus_mm.d15=Tarsus_mm, 
                 Weight_g.d15=Weight_g, Chao1.d15=Chao1, Shannon.d15=Shannon, PD.d15=PD, log.PD.d15=log.PD)

# merge dfs on bird id
paired.df <- merge(d8.df, d15.df, by="Bird.ID", all.x=T, all.y=T)

paired.df <- merge(paired.df, metadata[,c("Bird.ID", "Site","Nest","Treatment", "BroodSize.D8", "BroodSize.D15", "firstEggLayDate", "lastEggLayDate")], by="Bird.ID", all.x=T)
# rows being duplicated when bird ID occurs twice in metadata i.e. when D8 and D15 measured
# distinct drops duplicate rows
paired.df <- distinct(paired.df)
```

Rescale paired df
* not scaling weight at d15

```{r}
#dput(colnames(paired.df))
numeric.paired <- c( "Wing_mm.d8", "Tarsus_mm.d8", "Weight_g.d8", "Chao1.d8", 
                     "Shannon.d8",  "Wing_mm.d15", "Tarsus_mm.d15",
                     "Chao1.d15", "Shannon.d15","BroodSize.D8", "BroodSize.D15", "log.chao.d8", "PD.d8", "PD.d15", "log.PD.d8", "log.PD.d15") #"Weight_g.d15"

#paired.df.scaled <- paired.df
paired.df[,numeric.paired] <- scale(paired.df[,numeric.paired], scale = TRUE, center = TRUE)

```

Adding minimum scaled value of main explanatory variable (log.chao.d8) to that variable to make plots consistent with model output
```{r}
paired.df$log.chao.d8 <- paired.df$log.chao.d8 - min(paired.df$log.chao.d8, na.rm = T)
paired.df$Shannon.d8 <- paired.df$Shannon.d8 - min(paired.df$Shannon.d8, na.rm = T)
paired.df$log.PD.d8 <- paired.df$PD.d8 - min(paired.df$log.PD.d8, na.rm = T)
```

Repeat samples only
```{r}
repeat.df <- paired.df[!is.na(paired.df$Weight_g.d8) & !is.na(paired.df$Weight_g.d15),]
```

# Plot

No apparent difference except probiotic D8 look less variable than control D8

```{r}
ggplot(metadata.scaled, aes(Treatment, Weight_g, fill = Age.category)) +
  geom_boxplot() +
  theme_classic()
```

D8 nestlings' weight is less variable in treatment than control
```{r}
metadata.nst <- subset(x = metadata, subset = metadata$Age.category %in% c("D8"))
bartlett.test(metadata.nst$Weight_g, metadata.nst$Treatment)
```


Pretty normal
```{r}
hist(metadata.scaled$Weight_g)
```

# Analysis, weight

## All birds: treatment effect on weight

Using firstEggLayDate instead of Date
Using brood size at D8 because that will be closer to max brood size
No effect of treatment without diversity term
Tried chao and shannon, doesnt change results
Added logged chao1 following CC recommendation. Doesnt change results
```{r}
mod1 <- lmer(Weight_g ~ Treatment + Age.category.BDC + firstEggLayDate + BroodSize.D8 + log.chao +(1|Site/Nest/Bird.ID), data = metadata.scaled)
summary(mod1)
```

No interaction between treatment and age
no interaction between treatment and diversity
Added logged chao1 following CC recommendation. Result doesnt change qualitatively
```{r}
mod2 <- lmer(Weight_g ~ Treatment * log.chao + Treatment*Age.category.BDC + firstEggLayDate + BroodSize.D8 + (1|Site/Nest/Bird.ID), data = metadata.scaled)
summary(mod2)
```

```{r}
mod2.shannon <- lmer(Weight_g ~ Treatment * Shannon + Treatment*Age.category.BDC + firstEggLayDate + BroodSize.D8 + (1|Site/Nest/Bird.ID), data = metadata.scaled)
summary(mod2.shannon)
```
```{r}
mod2.PD <- lmer(Weight_g ~ Treatment * log.PD + Treatment*Age.category.BDC + firstEggLayDate + BroodSize.D8 + (1|Site/Nest/Bird.ID), data = metadata.scaled)
summary(mod2.PD)
```

## Nestling: treatment effect on weight

No effect of treatment on weight
No change when diversity term added
No interaction between diversity x treatment
No interaction between treatment x age
Adding log.chao, no qualitative change
```{r}
mod3 <- lmer(Weight_g ~ Treatment * Age.category + log.chao  + firstEggLayDate + BroodSize.D8 + (1|Site/Nest/Bird.ID), data = metadata.nestling)
summary(mod3)
```

No effect of treatment on weight
No change when diversity term added
```{r}
mod4 <- lmer(Weight_g ~ Treatment * Age.category + firstEggLayDate + BroodSize.D8 + Shannon + (1|Site/Nest/Bird.ID), data = metadata.nestling)
summary(mod4)
```

# Analysis, time-lag
## Time lag Chao

* significant interaction between treatment and chao, and including interaction makes chao significant
  * D8 chao diversity positively predicts D15 weight but this effect is much smaller in treatment birds
* changed to log.chao following CC recommendation, treatment is now significant
```{r}
mod5 <- lmer(Weight_g.d15 ~  Treatment * log.chao.d8 + Weight_g.d8 + firstEggLayDate + BroodSize.D8 +(1|Site/Nest), data = paired.df)
summary(mod5)
```

```{r}
simulationOutput <- simulateResiduals(fittedModel = mod5, n = 250)
plot(simulationOutput)
```

## time lag shannon
Only d8 weight predicts d15 weight whether or not shannon term included
* interaction between treatment and shannon is not significant
```{r}
mod6 <- lmer(Weight_g.d15 ~  Treatment * Shannon.d8 + Weight_g.d8 + firstEggLayDate + BroodSize.D8 +(1|Site/Nest), data = paired.df)
summary(mod6)
```

## time lag faiths PD

```{r}
mod7 <- lmer(Weight_g.d15 ~  Treatment * log.PD.d8 + Weight_g.d8 + firstEggLayDate + BroodSize.D8 +(1|Site/Nest), data = paired.df)
summary(mod7)
```

# save results
```{r}
save(mod1, mod2, mod2.shannon, mod3, mod4, mod5, mod6, file = "../Data/R_savedObjects/01_growthmodels.rda")
```

# Results tables

```{r}
chao.lm.coef <- summary(mod2)$coefficients # returns summary table for conditional average

rownames(chao.lm.coef) <- c("(Intercept)", "Treatment (L. kimchicus)", "log(Chao1)", "Age (D15-D8)", 
"Age (Adult-D15)", "First egg lay date", "Brood size (D8)", "Treatment (L. kimchicus) : log(Chao1)", "Treatment (L. kimchicus) : Age (D15)", "Treatment (L. kimchicus) : Age (Adult)")

#make df of lmer output
chao.lm.coef.df <- as_tibble(chao.lm.coef, rownames="Independent variables") %>%
  dplyr::rename("p"="Pr(>|t|)") %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(p=ifelse(p==0,"<0.001",p)) %>% 
  mutate(p=ifelse(p<=0.05,str_c(p," *"),p)) %>%
  mutate(p=ifelse(p>=0.05 & p<=0.06,str_c(p,"  ."),p))


# make df into kable
## this is in html, doesnt render in word doc # but can copy-paste
results.table <- kable(chao.lm.coef.df, format = "html", table.attr = "style = \"color: black;\"") %>%
  kableExtra::kable_styling(full_width = F) #%>%
  #save_kable("alpha-kable__________.png") # this line saves as .png in Reports/

results.table
```



```{r}
shannon.lm.coef <- summary(mod2.shannon)$coefficients # returns summary table for conditional average

rownames(shannon.lm.coef) <- c("(Intercept)", "Treatment (L. kimchicus)", "Shannon", "Age (D15-D8)",
"Age (Adult-D15)", "First egg lay date", "Brood size (D8)", "Treatment (L. kimchicus) : Shannon", "Treatment (L. kimchicus) : Age (D15)", "Treatment (L. kimchicus) : Age (Adult)")

#make df of lmer output
shannon.lm.coef.df <- as_tibble(shannon.lm.coef, rownames="Independent variables") %>%
  dplyr::rename("p"="Pr(>|t|)") %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(p=ifelse(p==0,"<0.001",p)) %>% 
  mutate(p=ifelse(p<=0.05,str_c(p," *"),p)) %>%
  mutate(p=ifelse(p>=0.05 & p<=0.06,str_c(p,"  ."),p))


# make df into kable
## this is in html, doesnt render in word doc # but can copy-paste
results.table <- kable(shannon.lm.coef.df, format = "html", table.attr = "style = \"color: black;\"") %>%
  kableExtra::kable_styling(full_width = F) #%>%
  #save_kable("alpha-kable__________.png") # this line saves as .png in Reports/

results.table
```

```{r}
PD.lm.coef <- summary(mod2.PD)$coefficients # returns summary table for conditional average

rownames(PD.lm.coef) <- c("(Intercept)", "Treatment (L. kimchicus)", "Faith's PD", "Age (D15-D8)",
"Age (Adult-D15)", "First egg lay date", "Brood size (D8)", "Treatment (L. kimchicus) : Faith's PD", "Treatment (L. kimchicus) : Age (D15)", "Treatment (L. kimchicus) : Age (Adult)")

#make df of lmer output
PD.lm.coef.df <- as_tibble(PD.lm.coef, rownames="Independent variables") %>%
  dplyr::rename("p"="Pr(>|t|)") %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(p=ifelse(p==0,"<0.001",p)) %>% 
  mutate(p=ifelse(p<=0.05,str_c(p," *"),p)) %>%
  mutate(p=ifelse(p>=0.05 & p<=0.06,str_c(p,"  ."),p))


# make df into kable
## this is in html, doesnt render in word doc # but can copy-paste
results.table <- kable(PD.lm.coef.df, format = "html", table.attr = "style = \"color: black;\"") %>%
  kableExtra::kable_styling(full_width = F) #%>%
  #save_kable("alpha-kable__________.png") # this line saves as .png in Reports/

results.table
```

```{r}
chao.int.lm.coef <- summary(mod5)$coefficients # returns summary table for conditional average

rownames(chao.int.lm.coef) <- c("(Intercept)", "Treatment (L. kimchicus)", "log(Chao1) (D8)", "Weight g (D8)", "First egg lay date", "Brood size (D8)", "Treatment (L. kimchicus) : Chao1 (D8)")

#make df of lmer output
chao.int.lm.coef.df <- as_tibble(chao.int.lm.coef, rownames="Independent variables") %>%
  dplyr::rename("p"="Pr(>|t|)") %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(p=ifelse(p==0,"<0.001",p)) %>% 
  mutate(p=ifelse(p<=0.05,str_c(p," *"),p)) %>%
  mutate(p=ifelse(p>=0.05 & p<=0.06,str_c(p,"  ."),p))


# make df into kable
## this is in html, dosnt render in word doc # but can copy-paste
results.table <- kable(chao.int.lm.coef.df, format = "html", table.attr = "style = \"color: black;\"") %>%
  kableExtra::kable_styling(full_width = F) #%>%
  #save_kable("alpha-kable__________.png") # this line saves as .png in Reports/

results.table
```


```{r}
shannon.int.lm.coef <- summary(mod6)$coefficients # returns summary table for conditional average

rownames(shannon.int.lm.coef) <- c("(Intercept)", "Treatment (L. kimchicus)", "Shannon (D8)", "Weight g (D8)", "First egg lay date", "Brood size (D8)", "Treatment (L. kimchicus) : Shannon (D8)")

#make df of lmer output
shannon.int.lm.coef.df <- as_tibble(shannon.int.lm.coef, rownames="Independent variables") %>%
  dplyr::rename("p"="Pr(>|t|)") %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(p=ifelse(p==0,"<0.001",p)) %>% 
  mutate(p=ifelse(p<=0.05,str_c(p," *"),p)) %>%
  mutate(p=ifelse(p>=0.05 & p<=0.06,str_c(p,"  ."),p))


# make df into kable
## this is in html, dosnt render in word doc # but can copy-paste
results.table <- kable(shannon.int.lm.coef.df, format = "html", table.attr = "style = \"color: black;\"") %>%
  kableExtra::kable_styling(full_width = F) #%>%
  #save_kable("alpha-kable__________.png") # this line saves as .png in Reports/

results.table
```

```{r}
PD.int.lm.coef <- summary(mod7)$coefficients # returns summary table for conditional average

rownames(PD.int.lm.coef) <- c("(Intercept)", "Treatment (L. kimchicus)", "Faith's PD (D8)", "Weight g (D8)", "First egg lay date", "Brood size (D8)", "Treatment (L. kimchicus) : Faith's PD (D8)")

#make df of lmer output
PD.int.lm.coef.df <- as_tibble(PD.int.lm.coef, rownames="Independent variables") %>%
  dplyr::rename("p"="Pr(>|t|)") %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(p=ifelse(p==0,"<0.001",p)) %>% 
  mutate(p=ifelse(p<=0.05,str_c(p," *"),p)) %>%
  mutate(p=ifelse(p>=0.05 & p<=0.06,str_c(p,"  ."),p))


# make df into kable
## this is in html, dosnt render in word doc # but can copy-paste
results.table <- kable(PD.int.lm.coef.df, format = "html", table.attr = "style = \"color: black;\"") %>%
  kableExtra::kable_styling(full_width = F) #%>%
  #save_kable("alpha-kable__________.png") # this line saves as .png in Reports/

results.table
```